---
title: "visualization_eda"
author: "Leighanne Wang"
date: "10/6/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)

```

## Load weather data

```{r weather_df}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Different ways to create a Scatterplot

Create first scatterplot:
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

New approach, same plot:
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Save and edit plot object
```{r}
#create a plot
weather_plot =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

#add to the plot
weather_plot + geom_point()
```

## Advanced Scatterplot

color
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()
```

smoothing
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```

'aes' placement
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name))

# ends up being the same as the first graph
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + #having aes here is a global command
  geom_point(aes(color = name)) + #color only gets applied to scatterplot not to the entire thing like geom smooth
  geom_smooth()
```

faceting
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) + #alpha changes the transparency for the dots 
  geom_smooth(se = FALSE, size = 2) + # size changes the size of the line
  facet_grid(. ~ name)
  #. means don't create rows, create columns by variable name 
```

combine elements to create new plot
```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## Small notes

how many geoms have to exist? as many as you want!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

you can use a neat geom

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density2d() +
  geom_point(alpha = .3)
```


## Univariate plots

####Histograms

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

adding color

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + #fill instead of color to make the whole bar that color
  geom_histogram(position = "dodge") #position dodge is to move it so that they're different location
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + #fill instead of color to make the whole bar that color
  geom_histogram()
  facet_grid(. ~ name)
```

new geometry
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```

#### boxplots

```{r}
weather_df %>% 
  ggplot(aes(y = tmin)) +
  geom_boxplot()
```

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

#### trendy plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
  stat_summary()
```

ridge plots
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Save and Embed

save a scatterplot

```{r}
weather_plot =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)
```

embed
```{r, fig.width=6, fig.height=3}
weather_plot
```

```{r, fig.width=12, fig.height=6}
weather_plot
```



















